---
title: "Power"
output: html_document
date: '2022-12-17'
---

```{r, setup, include=F}
# Clear workspace of all objects and unload all extra (non-base) packages
rm(list = ls(all = TRUE))
if (!is.null(sessionInfo()$otherPkgs)) {
  res <- suppressWarnings(
    lapply(paste('package:', names(sessionInfo()$otherPkgs), sep=""),
           detach, character.only=TRUE, unload=TRUE, force=TRUE))
}

knitr::opts_chunk$set(cache=F, cache.comments = F, message = F, warning = F, tidy.opts=list(width.cutoff=60), tidy=TRUE)  

pacman::p_load(tidyverse, dplyr,  pwr #,knitr, powerSurvEpi
               )

# ggplot settings
theme_set(theme_bw())
theme_update(legend.position = "bottom")

set.seed(1)

```

# values 

```{r}
(n_tot <- 2310 + 2428)

# use?
n_child <- 1670
n_adult <- 3068


pwr_lvl <- 0.80
sig_lvl <- 0.05

# Ospina 2022 2013-2014 overall values
ospina_mean <- 0.411
ospina_upper <- 0.450

# upper_CI = mean + 1.96*SE
# SE = (upper_CI - mean)/1.96

# SE = SD/sqrt(N)
# SD = sqrt(N)*SE
# SD = sqrt(n)*[(upper_CI - mean)/1.96]
 
(sd_pooled <- sqrt(n_tot)*(ospina_upper - ospina_mean)/1.96) #1.37 ug/L

```


# t-test

could replace n_tot with n_adults or n_children

```{r}


print("effect size we will be able to detect given a sample size, power, and significance level")
pwr1 <- pwr.t.test(n=n_tot/2, power = pwr_lvl, sig.level = sig_lvl,
           type="two.sample", alternative="two.sided")

# d = (u1 - u2)/sd_pooled)
# u1 - u2 = d*sd_pooled
pwr1$d*sd_pooled

```

# effect sizes detected in the literature



## Ospina 2022 

fasting > 8 hr vs ≤ 8 hr  (Table 2), ug/L     
* children: `r 0.51-0.43 # 0.08` 
* adults: `r 0.44-0.37 # 0.07`   
* overall: 0.10 (Table S1,ug/g creatinine)   

overall (table S1), (ug/g creatinine)
* cereal consumption: `r 0.517 - 0.422 # 0.09`     
* soft drink consumption:  `r 0.472 - 0.405 # 0.067`   
* season of sample collected: `r 0.478 - 0.401 # 0.077`    
* used products to kill weeds: `r 0.473 - 0.442 # 0.031`
 
## Curl manuscript in proress
* Table 3, near field vs far from field GM difference is: `r 0.19-0.16 #0.03`


## Gillezeau  2019 review     
note that these studies used different methods and diferent LODs 

* general population avg: 0.16-7.6    
* occupationally exposed individuals: 0.26-73.5   

the minimum difference is about 0.10, close to the effect size we expect to be able to detect

## Curwin ___ farm vs non-farm urinary gly in Iowa
* Table 2 non-farm vs farm: 0.3 (mother), 0.5 (father), 0.7 (child)

* we expect to see higher variability in NHANES since estimates will be from places throughout the US

# Power Language (not used in the end)

"Using a two-sided, two-sample t-test for a total sample size of 4,738, we calculated having 80% power at a 0.05 significance level to detect an effect size of 0.11 ug/L given a pooled SD similar to the 2013-2014 NHANES of 1.37 ug/L (Ospina et al. 2022)."

Note: This estimate: 1) does not stratify by children vs adults; 2) assumes that comparison groups are the same size

